Synthetic data

### 2022-04-29 dev/work/synthetic_bayesian_ar.qmd
aies

```{julia}
include("../src/load.jl")
using AlgorithmicRecourse
using CSV;
using DataFrames;
using LinearAlgebra;
using Distributions;
output_folder = "output/synthetic"
www_folder = "www/synthetic"
using Base.Filesystem: joinpath
using Images, FileIO
```

@@

```{julia}
using AlgorithmicRecourse.Models: BayesianLogisticModel
𝑴 = BayesianLogisticModel(reshape(model.μ,1,length(model.μ)), model.Σ); # repare for usage with AlgorithmicRecourse.jl
x̅ = [50.0, 60.0]; # factual
using Random
Random.seed!(1234)
y̅ = round(AlgorithmicRecourse.Models.probs(𝑴, x̅)[1])
target = ifelse(y̅==1.0,0.0,1.0) # opposite label as target
γ = ifelse(target==1.0,0.9,0.1); # desired threshold based on target
```

>>>>

```{julia}
include("../src/load.jl")
using CounterfactualExplanations
using CSV;
using DataFrames;
using LinearAlgebra;
using Distributions;
output_folder = "output/synthetic"
www_folder = "www/synthetic"
using Base.Filesystem: joinpath
using Images, FileIO
```

@@

```{julia}
using CounterfactualExplanations.Models: BayesianLogisticModel
𝑴 = BayesianLogisticModel(reshape(model.μ,1,length(model.μ)), model.Σ); # repare for usage with CounterfactualExplanations.jl
x̅ = [50.0, 60.0]; # factual
using Random
Random.seed!(1234)
y̅ = round(CounterfactualExplanations.Models.probs(𝑴, x̅)[1])
target = ifelse(y̅==1.0,0.0,1.0) # opposite label as target
γ = ifelse(target==1.0,0.9,0.1); # desired threshold based on target
```

### 2022-04-29 dev/work/synthetic_endogenous_shifts.qmd
aies

---
jupyter: julia-1.6
---

# Endogenous model shifts

@@

```{julia}
include("dev/utils.jl")
using CounterfactualExplanations, CSV, DataFrames, LinearAlgebra, Distributions, Images, FileIO, Plots, PlotThemes
using Base.Filesystem: joinpath
theme(:wong)
using Logging
disable_logging(Logging.Info)
output_folder = "output/synthetic"
www_folder = "www/synthetic"
```

## Single-period

### Training the classifier

We start by training a simple linear classifier: Bayesian Logisitic Regression.

```{julia}
df = CSV.read("../data/cats_dogs_large.csv", DataFrame);
y = df[:,:y];
N = length(y);
X = Matrix(df[:,Not(:y)]);
model = Models.bayes_logreg(X,y); # get MAP estimates
w = model.μ;
```

```{julia}
# Plot data points.
plt = plot()
using Main.Experiments: plot_data!
plot_data!(plt,X,y)
```

```{julia}
using AlgorithmicRecourse.Models: BayesianLogisticModel, LogisticModel
𝑴 = BayesianLogisticModel(reshape(model.μ,1,length(model.μ)), model.Σ) # fully Bayesian model
𝑴_plugin = LogisticModel(reshape(model.μ[2:3],1,2), [model.μ[1]]); # Plugin estimate (non-Bayesian)
models = (Bayesian=𝑴, Plugin=𝑴_plugin);
```

```{julia}
using Main.Experiments: plot_contour
p1 = plot_contour(X,y,𝑴,title="Bayesian")
p2 = plot_contour(X,y,𝑴_plugin,title="Plugin")
plot(p1,p2,size=(1000,400))
```

### Generating recourse

Now we will generate and implement algorithmic recourse for a subset of cats.

```{julia}
generators = (Bayesian=GreedyGenerator(1,500,:logitbinarycrossentropy, nothing), Plugin=GenericGenerator(0.1,0.1,1e-5,:logitbinarycrossentropy,nothing));
proportion = 0.2
using StatsBase
cats = findall(df.y.==0);
chosen_individuals = StatsBase.sample(cats,Int(round(proportion * length(cats))),replace=false);
target = 1.0
γ = 0.90;
```

```{julia}
function generate_counterfactuals(generator, 𝑴, target, γ, chosen_individuals; X=X, y=y)
    X̲ = copy(X)
    y̲ = copy(y)
    for i in chosen_individuals
        x̅ = X[i,:]
        recourse = generate_recourse(generator, x̅, 𝑴, target, γ);
        X̲[i,:] = recourse.x̲
        y̲[i] = recourse.y̲
    end
    return X̲, y̲
end
```

```{julia}
plt = []
for i in 1:length(generators)
    X̲, y̲ = generate_counterfactuals(generators[i], models[i], target, γ, chosen_individuals)
    pltᵢ = plot_contour(X̲,y̲,models[i],title=keys(generators)[i])
    plt = vcat(plt, pltᵢ)
end
plot(plt..., size = (1000,400))
```

### Retraining classifier

Finally, we will retrain the classifier (model shift) and plot the resulting new decision boundary and posterior predictive distribution.

```{julia}
plt = []
for i in 1:length(generators)
    X̲, y̲ = generate_counterfactuals(generators[i], models[i], target, γ, chosen_individuals)
    data = Models.prepare_data(X̲', y̲')
    𝑴 = Models.retrain(models[i], data)
    pltᵢ = plot_contour(X̲,y̲,𝑴,title=keys(generators)[i])
    plt = vcat(plt, pltᵢ)
end
plot(plt..., size = (1000,400))
```

## Dynamic

```{julia}
# Variables:
μ = [0.01,0.1,0.25]
γ = [0.50,0.75,0.9]
grid_ = Experiments.GridVariables(μ, γ)
n_rounds = 10
n_folds = 5
target = 1.0
T = 1000;
# Experiments:
experiment = Experiments.Experiment(X',y',models[:Bayesian],target,grid_,n_rounds)
experiment_plugin = Experiments.Experiment(X',y',models[:Plugin],target,grid_,n_rounds)
experiments = (experiment, experiment_plugin);
```

```{julia}
run = false
if run
    using Random
    Random.seed!(1234)
    for j in 1:length(generators)
        outcome, path = Experiments.run_experiment(experiments[j], generators[j], n_folds, T=T, store_path=true)
        Experiments.save_path(joinpath(output_folder, string(keys(generators)[j])),path)
        CSV.write(joinpath(output_folder, string(keys(generators)[j]) * ".csv"), DataFrame(outcome))
    end
end
```

```{julia}
run = false
if run
    using Main.Experiments: plot_path
    paths = [Experiments.load_path(joinpath(output_folder, string(keys(generators)[j]))) for j ∈ 1:length(generators)];
    anim = Experiments.plot_path(paths[1],Γ=γ,𝙈=μ)
    gif(anim,joinpath(www_folder,"bayesian.gif"), fps=2)
    anim = Experiments.plot_path(paths[2],Γ=γ,𝙈=μ)
    gif(anim,joinpath(www_folder,"plugin.gif"), fps=2)
end
```

@@

```{julia}
using Main.Experiments: prepare_results, plot_results
validity = DataFrame()
cost = DataFrame()
files = (Bayesian=joinpath(output_folder, "Bayesian.csv"), Plugin=joinpath(output_folder, "Plugin.csv"))
results = map(file -> CSV.read(file, DataFrame), files)
p_val, p_cost = plot_results(results);
```

@@

  gif(anim, joinpath(www_folder, "single_nn.gif"), fps=10)

  BSON.@save joinpath(output_folder, "nn.bson") nn

@@

```{julia}
using BSON: @load
@load joinpath(output_folder, "nn.bson") nn
𝑴ₙₙ = Models.FittedNeuralNet(nn, opt, loss);
```

@@

```{julia}
opt = ADAM()
loss_type = :logitbinarycrossentropy
run = false
if run
    K = 50
    ensemble = Models.build_ensemble(K,kw=(input_dim=size(X')[1], n_hidden=100));
    ensemble, anim = Models.forward(ensemble, data, opt, n_epochs=10, plot_every=10, loss_type=loss_type, τ=1.1); # fit the ensemble
    Models.save_ensemble(ensemble, root=joinpath(output_folder, "ensemble")) # save to disk
    gif(anim, joinpath(www_folder, "ensemble_loss.gif"), fps=25);
end
```

```{julia}
ensemble = Models.load_ensemble(root=joinpath(output_folder, "ensemble"))
𝑴 = Models.FittedEnsemble(ensemble, opt, loss_type)
models = (Bayesian=𝑴, Plugin=𝑴ₙₙ);
```

@@

```{julia}
run = true
if run
    using Random
    Random.seed!(1234)
    for j in 1:length(generators)
        outcome, path = Experiments.run_experiment(experiments[j], generators[j], n_folds, T=T, store_path=true)
        Experiments.save_path(joinpath(output_folder, string(keys(generators)[j])),path)
        CSV.write(joinpath(output_folder, string(keys(generators)[j]) * "_deep.csv"), DataFrame(outcome))
    end
end
```

>>>>

---
title: Synthetic data
jupyter: julia-1.6
---

@@

```{julia}
include("src/load.jl")
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, BayesLaplace, LinearAlgebra
using Base.Filesystem: joinpath
theme(:wong)
using Logging
disable_logging(Logging.Info)
output_path = "dev/work/output/synthetic"
www_path = "dev/work/www/synthetic"
```

## Classifiers

```{julia}
# Number of points to generate.
N = 100
using CounterfactualExplanations.Data
xs, ys = Data.toy_data_linear(N)
X = hcat(xs...) # bring into tabular format
data = zip(xs,ys)
```

### Logistic Regression

```{julia}
nn = Chain(Dense(2,1))
λ = 0.5
sqnorm(x) = sum(abs2, x)
weight_regularization(λ=λ) = 1/2 * λ^2 * sum(sqnorm, Flux.params(nn))
loss(x, y) = Flux.Losses.logitbinarycrossentropy(nn(x), y) + weight_regularization();
```

```{julia}
using Flux.Optimise: update!, ADAM
opt = ADAM()
epochs = 50
function _forward_nn(nn, data, epochs)
    for epoch = 1:epochs
        for d in data
            gs = gradient(params(nn)) do
            l = loss(d...)
            end
            update!(opt, params(nn), gs)
        end
    end
    return nn
end
_forward_nn(nn, data, epochs)
```

```{julia}
using CounterfactualExplanations.Models
import CounterfactualExplanations.Models: logits, probs # import functions in order to extend
import Main.Models: retrain

# Step 1)
struct NeuralNetwork <: Models.AbstractFittedModel
    model::Any
end

# Step 2)
logits(M::NeuralNetwork, X::AbstractArray) = M.model(X)
probs(M::NeuralNetwork, X::AbstractArray)= σ.(logits(M, X))

# Step 3)
function retrain(M::NeuralNetwork, data; n_epochs=10, τ=1.0) 
    nn = M.model
    nn = _forward_nn(nn, data, n_epochs)
    M = NeuralNetwork(nn)
    return M
end

M = NeuralNetwork(nn)
```

```{julia}
#| echo: false
# Plot the posterior distribution with a contour plot.
plt = plot_contour(X',ys,M);
savefig(plt, joinpath(www_path, "nn_contour.png"))
```

![](www/nn_contour.png)

### Laplace Redux

```{julia}
la = Laplace(nn, λ=λ)
BayesLaplace.fit!(la, data)
```

```{julia}
# Step 1)
struct LaplaceNeuralNetwork <: Models.AbstractFittedModel
    la::BayesLaplace.Laplace
end

# Step 2)
logits(M::LaplaceNeuralNetwork, X::AbstractArray) = M.la.model(X)
probs(M::LaplaceNeuralNetwork, X::AbstractArray)= BayesLaplace.predict(M.la, X)

# Step 3)
function retrain(M::LaplaceNeuralNetwork, data; n_epochs=10, τ=1.0) 
    nn = M.la.model
    nn = _forward_nn(nn, data, n_epochs)
    la = Laplace(nn, λ=λ)
    BayesLaplace.fit!(la, data)
    M = LaplaceNeuralNetwork(la)
    return M
end

Mᴸ = LaplaceNeuralNetwork(la)
```

```{julia}
#| echo: false
# Plot the posterior distribution with a contour plot.
plt = plot_contour(X',ys,Mᴸ);
savefig(plt, joinpath(www_path, "la_contour.png"))
```

![](www/la_contour.png)

## Experiment

```{julia}
# Models:
models = (Bayesian=Mᴸ, Plugin=M)

# Generators:
generators = (Bayesian=GreedyGenerator(), Plugin=GenericGenerator())

# Variables:
μ = [0.01,0.1,0.25]
γ = [0.50,0.75,0.9]
grid_ = Experiments.GridVariables(μ, γ)
n_rounds = 10
n_folds = 5
target = 1.0
T = 1000
```

```{julia}
# Experiments:
experiment = Experiments.Experiment(X,ys,models[:Bayesian],target,grid_,n_rounds)
experiment_plugin = Experiments.Experiment(X,ys,models[:Plugin],target,grid_,n_rounds)
experiments = (experiment, experiment_plugin)
```

```{julia}
#| eval: false
using Random, CSV, DataFrames
Random.seed!(1234)
for j in 1:length(generators)
    outcome, path = Experiments.run_experiment(experiments[j], generators[j], n_folds, T=T, store_path=true)
    Experiments.save_path(joinpath(output_path, string(keys(generators)[j])),path)
    CSV.write(joinpath(output_path, string(keys(generators)[j]) * ".csv"), DataFrame(outcome))
end
```

```{julia}
#| eval: false
using Main.Experiments: plot_path
paths = [Experiments.load_path(joinpath(output_path, string(keys(generators)[j]))) for j ∈ 1:length(generators)];
anim = Experiments.plot_path(paths[1],Γ=γ,𝙈=μ)
gif(anim,joinpath(www_path,"bayesian.gif"), fps=2)
anim = Experiments.plot_path(paths[2],Γ=γ,𝙈=μ)
gif(anim,joinpath(www_path,"plugin.gif"), fps=2)
```

@@

```{julia}
using Main.Experiments: prepare_results, plot_results
validity = DataFrame()
cost = DataFrame()
files = (Bayesian=joinpath(output_path, "Bayesian.csv"), Plugin=joinpath(output_path, "Plugin.csv"))
results = map(file -> CSV.read(file, DataFrame), files)
p_val, p_cost = plot_results(results);
```

@@

  gif(anim, joinpath(www_path
, "single_nn.gif"), fps=10)

  BSON.@save joinpath(output_path
, "nn.bson") nn

@@

```{julia}
using BSON: @load
@load joinpath(output_path, "nn.bson") nn
Mₙₙ = Models.FittedNeuralNet(nn, opt, loss);
```

@@

```{julia}
opt = ADAM()
loss_type = :logitbinarycrossentropy
run = false
if run
    K = 50
    ensemble = Models.build_ensemble(K,kw=(input_dim=size(X')[1], n_hidden=100));
    ensemble, anim = Models.forward(ensemble, data, opt, n_epochs=10, plot_every=10, loss_type=loss_type, τ=1.1); # fit the ensemble
    Models.save_ensemble(ensemble, root=joinpath(output_path
, "ensemble")) # save to disk
    gif(anim, joinpath(www_path
, "ensemble_loss.gif"), fps=25);
end
```

```{julia}
ensemble = Models.load_ensemble(root=joinpath(output_path, "ensemble"))
M = Models.FittedEnsemble(ensemble, opt, loss_type)
models = (Bayesian=M, Plugin=Mₙₙ);
```

@@

```{julia}
run = true
if run
    using Random
    Random.seed!(1234)
    for j in 1:length(generators)
        outcome, path = Experiments.run_experiment(experiments[j], generators[j], n_folds, T=T, store_path=true)
        Experiments.save_path(joinpath(output_path
    , string(keys(generators)[j])),path)
        CSV.write(joinpath(output_path
    , string(keys(generators)[j]) * "_deep.csv"), DataFrame(outcome))
    end
end
```
